---
title: "data_mgmt"
author: "Wendel Raymond"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r}
## Library ##
library(tidyverse)
```


## SJI pilot eelgrass sensor deployment data managment
Import raw data and prepare for plotting and analysis.

### Data
Including sensor deployment metadata and sensor data files themselves.
```{r}
## Deployments metadata ##
deploy <- read.csv("sensor_deployments.csv", header  = TRUE, stringsAsFactors = FALSE)

## PAR Calibration coefficients ##
par.coefs <- read.csv('Calibration_2019_12.csv') %>%
  select(sensor, medium, estimate) %>%
  filter(medium == "water") %>% 
  mutate(sensor_sn = as.character(sensor))

## Sensor data ##
ImportAll.cvs <- function(directory) {
  # Get a list of all .csv files in the specified directory
  csv_files <- list.files(path = directory, pattern = "\\.csv$", full.names = TRUE)
  
  # Initialize an empty list to store the data frames
  data_frames <- list()
  
  # Loop through each .csv file and read it into a data frame
  for (file in csv_files) {
    data <- read.csv(file, header = FALSE)
    data_frames[[basename(file)]] <- data
  }
  
  # Return a list of data frames
  return(data_frames)
}

sensors.raw <- ImportAll.cvs("data/")
data.names <- names(sensors.raw)
```

### Deployment metadata
Inspect deployments data to extract identifying information.
```{r}
deploy.sji <- deploy %>% 
  filter(site == "SJI") %>% 
  filter(launch.date == "5/23/2023")
```

### By sensor type data management
For each sensor type, select the appropriate data files and clean up format to long with rows and observations and columns as timestamp and values.

#### Temperature
From dedicated temperature sensor only

```{r}
## Select temperature metadata ##
temp.sns <- deploy.sji %>% 
  filter(type == "temp") %>% 
  select(sensor, site, subsite)

temp.sns.vc <- temp.sns$sensor

## Loop over sensors to create master temperature dataframe ##
temp.dat <- data.frame()

for(i in 1:length(temp.sns.vc)){
  dat.name <- paste("SJI", temp.sns[i, 3], "2023_02", temp.sns[i, 1], sep = "_")
  dat.name <- paste(dat.name, "csv", sep = ".")
  dat <- sensors.raw[[dat.name]]
  dat.new <- data.frame(site = substr(dat.name, start = 1, stop = 5),
                       sensor_sn = substr(dat.name, start = 15, stop = 22),
                       time = dat[5:nrow(dat), 2],
                       temp_c = dat[5:nrow(dat), 3])
  
  temp.dat <- rbind(temp.dat, dat.new)
}

temp.dat <- temp.dat %>% 
  mutate(time = as.POSIXct(time, format = "%m/%d/%Y %H:%M:%S")) %>% 
  mutate(temp_c = as.numeric(temp_c))
```

#### Conductivity
This sensor records high and low conductivity and temperature. 

```{r}
## Select conductivity metadata ##
cond.sns <- deploy.sji %>% 
  filter(type == "salinity") %>% 
  select(sensor, site, subsite)

cond.sns.vc <- cond.sns$sensor

## Loop over sensors to create master conductivity dataframe ##
cond.dat <- data.frame()

for(i in 1:length(cond.sns.vc)){
  dat.name <- paste("SJI", cond.sns[i, 3], "2023_02", cond.sns[i, 1], sep = "_")
  dat.name <- paste(dat.name, "csv", sep = ".")
  dat <- sensors.raw[[dat.name]]
  dat.new <- data.frame(site = substr(dat.name, start = 1, stop = 5),
                       sensor_sn = substr(dat.name, start = 15, stop = 22),
                       time = dat[3:nrow(dat), 2],
                       spc_cond_uScm = dat[3:nrow(dat), 6],
                       sal_ppt = dat[3:nrow(dat), 7],
                       temp_c = dat[3:nrow(dat), 5],
                       lo_cond_uScm = dat[3:nrow(dat), 3],
                       hi_cond_uScm = dat[3:nrow(dat), 4])
  
  cond.dat <- rbind(cond.dat, dat.new)
}

cond.dat <- cond.dat %>% 
  mutate(time = strptime(time, format = "%m/%d/%y %I:%M:%S %p")) %>% 
  mutate(time = as.POSIXct(time, format = "%Y-%m-%d %H:%M:%S")) %>%  
  mutate(spc_cond_uScm = as.numeric(spc_cond_uScm)) %>%
  mutate(sal_ppt = as.numeric(sal_ppt)) %>%
  mutate(temp_c = as.numeric(temp_c)) %>% 
  mutate(lo_cond_uScm = as.numeric(lo_cond_uScm)) %>% 
  mutate(hi_cond_uScm = as.numeric(hi_cond_uScm))
```

#### PAR
Note these data are not (un-) calibrated. Calibration should occur elsewhere.

```{r}
## Select PAR metadata ##
par.sns <- deploy.sji %>% 
  filter(type == "par") %>% 
  select(sensor, site, subsite)

par.sns.vc <- par.sns$sensor

## Loop over sensors to create master par dataframe ##
par.dat <- data.frame()

for(i in 1:length(par.sns.vc)){
  dat.name <- paste("SJI", par.sns[i, 3], "2023_02", par.sns[i, 1], sep = "_")
  dat.name <- paste(dat.name, "csv", sep = ".")
  dat <- sensors.raw[[dat.name]]
  dat.new <- data.frame(site = substr(dat.name, start = 1, stop = 5),
                       sensor_sn = substr(dat.name, start = 15, stop = 18),
                       time = paste(dat[7:nrow(dat), 2], dat[7:nrow(dat), 3], sep = " "),
                       par_uncal = dat[7:nrow(dat), 4])
  
  par.dat <- rbind(par.dat, dat.new)
}

par.dat <- par.dat %>% 
  mutate(time = as.POSIXct(time, format = "%d/%m/%Y %H:%M:%S")) %>% 
  mutate(par_uncal = as.numeric(par_uncal))
```

#### Depth
Note these data are not direct readings of depth, rather measures of absolute pressure in kPa. These data will need to be calibrated. Also, these sensor also report temperature.
```{r}
## Select depth metadata ##
depth.sns <- deploy.sji %>% 
  filter(type == "depth") %>% 
  select(sensor, site, subsite)

depth.sns.vc <- par.sns$sensor

## Loop over sensors to create master depth dataframe ##
depth.dat <- data.frame()

for(i in 1:length(depth.sns.vc)){
  dat.name <- paste("SJI", depth.sns[i, 3], "2023_02", depth.sns[i, 1], sep = "_")
  dat.name <- paste(dat.name, "csv", sep = ".")
  dat <- sensors.raw[[dat.name]]
  dat.new <- data.frame(site = substr(dat.name, start = 1, stop = 5),
                       sensor_sn = substr(dat.name, start = 15, stop = 22),
                       time = dat[3:nrow(dat), 2],
                       abs_pres_kPa = dat[3:nrow(dat), 3],
                       temp_c = dat[3:nrow(dat), 4])
  
  depth.dat <- rbind(depth.dat, dat.new)
}

depth.dat <- depth.dat %>% 
  mutate(time = strptime(time, format = "%m/%d/%y %I:%M:%S %p")) %>% 
  mutate(time = as.POSIXct(time, format = "%Y-%m-%d %H:%M:%S")) %>% 
  mutate(abs_pres_kPa = as.numeric(abs_pres_kPa)) %>% 
  mutate(temp_c = as.numeric(temp_c))
```

### Calibration and correction
Calibrate and correct data as needed. 

#### PAR
Use calibration coefficients to correct raw data
```{r}
par.dat <- left_join(par.dat, par.coefs[, 3:4], by = "sensor_sn") %>%
  mutate(estimate = if_else(is.na(estimate), 1, estimate), # If a sensor has no calibration event, use 1 as coefficient
         calib = par_uncal * estimate,
         par_cal = par_uncal * estimate * 900 *0.000001)
```

#### Depth

#### Conductivity

### Exploratory plots

#### Add some metadata
Site names

```{r}
## Temperature ##
temp.dat <- temp.dat %>%
  mutate(site.name = case_when(
         site == "SJI_1" ~ "Orcas E",
         site == "SJI_2" ~ "Orcas W",
         site == "SJI_3" ~ "Picnic Cove",
         site == "SJI_4" ~ "Reef Net Bay",
         site == "SJI_5" ~ "Jackson Beach",
         site == "SJI_6" ~ "4th of July",
         site == "SJI_7" ~ "Westcott Bay",
         site == "SJI_8" ~"Mitchell Bay")
         )

temp.dat$site.name <- factor(temp.dat$site.name, levels = c("Orcas E", "Orcas W",
                                                            "Mitchell Bay", "Westcott Bay",
                                                            "Reef Net Bay", "Picnic Cove",
                                                            "4th of July", "Jackson Beach"))

## Salinity ##
cond.dat <- cond.dat %>%
  mutate(site.name = case_when(
         site == "SJI_1" ~ "Orcas E",
         site == "SJI_2" ~ "Orcas W",
         site == "SJI_3" ~ "Picnic Cove",
         site == "SJI_4" ~ "Reef Net Bay",
         site == "SJI_5" ~ "Jackson Beach",
         site == "SJI_6" ~ "4th of July",
         site == "SJI_7" ~ "Westcott Bay",
         site == "SJI_8" ~"Mitchell Bay")
         )

cond.dat$site.name <- factor(cond.dat$site.name, levels = c("Orcas E", "Orcas W",
                                                            "Mitchell Bay", "Westcott Bay",
                                                            "Reef Net Bay", "Picnic Cove",
                                                            "4th of July", "Jackson Beach"))
```


#### Temperature
```{r}
ggplot(temp.dat) + 
  geom_point(aes(x = time, y = temp_c)) + 
  facet_wrap(~site.name, ncol = 2, nrow = 4)
```

#### Conductivity and Salinity

```{r}
ggplot(cond.dat) + 
  geom_point(aes(x = time, y = sal_ppt)) + 
  facet_wrap(~site.name, ncol = 2, nrow = 4)
```

