---
title: "data_mgmt"
author: "Wendel Raymond"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r}
## Library ##
library(tidyverse)

theme_set(theme_classic())
```


## SJI pilot eelgrass sensor deployment data managment
Import raw data and prepare for plotting and analysis.

### Data
Including sensor deployment metadata and sensor data files themselves.
```{r}
## Deployments metadata ##
deploy <- read.csv("sensor_deployments.csv", header  = TRUE, stringsAsFactors = FALSE)

## PAR Calibration coefficients ##
par.coefs <- read.csv('Calibration_2019_12.csv') %>%
  select(sensor, medium, estimate) %>%
  filter(medium == "water") %>% 
  mutate(sensor_sn = as.character(sensor))

## Sensor data ##
ImportAll.cvs <- function(directory) {
  # Get a list of all .csv files in the specified directory
  csv_files <- list.files(path = directory, pattern = "\\.csv$", full.names = TRUE)
  
  # Initialize an empty list to store the data frames
  data_frames <- list()
  
  # Loop through each .csv file and read it into a data frame
  for (file in csv_files) {
    data <- read.csv(file, header = FALSE)
    data_frames[[basename(file)]] <- data
  }
  
  # Return a list of data frames
  return(data_frames)
}

sensors.raw <- ImportAll.cvs("data/")
data.names <- names(sensors.raw)
```

### Deployment metadata
Inspect deployments data to extract identifying information.
```{r}
deploy.sji <- deploy %>% 
  filter(site == "SJI") %>% 
  filter(launch.date == "5/23/2023")
```

### By sensor type data management
For each sensor type, select the appropriate data files and clean up format to long with rows and observations and columns as timestamp and values.

```{r}
site.names <- data.frame(
  site = c("SJI_1", "SJI_2", "SJI_3", "SJI_4", "SJI_5", "SJI_6", "SJI_7", "SJI_8"),
  site_name = c("Ocras East", "Orcas West", "Picnic Cove", "Reef Net Bay", "Jackson Beach", "4th of July", "Westcott Bay", "Mitchell Bay")
)
```


#### Temperature
From dedicated temperature sensor only

```{r}
## Select temperature metadata ##
temp.sns <- deploy.sji %>% 
  filter(type == "temp") %>% 
  select(sensor, site, subsite)

temp.sns.vc <- temp.sns$sensor

## Loop over sensors to create master temperature dataframe ##
temp.dat <- data.frame()

for(i in 1:length(temp.sns.vc)){
  dat.name <- paste("SJI", temp.sns[i, 3], "2023_02", temp.sns[i, 1], sep = "_")
  dat.name <- paste(dat.name, "csv", sep = ".")
  dat <- sensors.raw[[dat.name]]
  dat.new <- data.frame(site = substr(dat.name, start = 1, stop = 5),
                       sensor_sn = substr(dat.name, start = 15, stop = 22),
                       time = dat[5:nrow(dat), 2],
                       temp_c = dat[5:nrow(dat), 3])
  
  temp.dat <- rbind(temp.dat, dat.new)
}

temp.dat <- temp.dat %>% 
  mutate(time = as.POSIXct(time, format = "%m/%d/%Y %H:%M:%S")) %>% 
  mutate(temp_c = as.numeric(temp_c))

temp.dat <- left_join(temp.dat, site.names)
```

#### Conductivity
This sensor records high and low conductivity and temperature. 

```{r}
## Select conductivity metadata ##
cond.sns <- deploy.sji %>% 
  filter(type == "salinity") %>% 
  select(sensor, site, subsite)

cond.sns.vc <- cond.sns$sensor

## Loop over sensors to create master conductivity dataframe ##
cond.dat <- data.frame()

for(i in 1:length(cond.sns.vc)){
  dat.name <- paste("SJI", cond.sns[i, 3], "2023_02", cond.sns[i, 1], "corr", sep = "_")
  dat.name <- paste(dat.name, "csv", sep = ".")
  dat <- sensors.raw[[dat.name]]
  dat.new <- data.frame(site = substr(dat.name, start = 1, stop = 5),
                       sensor_sn = substr(dat.name, start = 15, stop = 22),
                       time = dat[3:nrow(dat), 2],
                       spc_cond_uScm = dat[3:nrow(dat), 6],
                       sal_ppt = dat[3:nrow(dat), 7],
                       temp_c = dat[3:nrow(dat), 5],
                       lo_cond_uScm = dat[3:nrow(dat), 3],
                       hi_cond_uScm = dat[3:nrow(dat), 4])
  
  cond.dat <- rbind(cond.dat, dat.new)
}

cond.dat <- cond.dat %>% 
  mutate(time = strptime(time, format = "%m/%d/%y %I:%M:%S %p")) %>% 
  mutate(time = as.POSIXct(time, format = "%Y-%m-%d %H:%M:%S")) %>%  
  mutate(spc_cond_uScm = as.numeric(spc_cond_uScm)) %>%
  mutate(sal_ppt = as.numeric(sal_ppt)) %>%
  mutate(temp_c = as.numeric(temp_c)) %>% 
  mutate(lo_cond_uScm = as.numeric(lo_cond_uScm)) %>% 
  mutate(hi_cond_uScm = as.numeric(hi_cond_uScm))

cond.dat <- left_join(cond.dat, site.names)
```

#### PAR
Note these data are not (un-) calibrated. Calibration should occur elsewhere.

```{r}
## Select PAR metadata ##
par.sns <- deploy.sji %>% 
  filter(type == "par") %>% 
  select(sensor, site, subsite)

par.sns.vc <- par.sns$sensor

## Loop over sensors to create master par dataframe ##
par.dat <- data.frame()

for(i in 1:length(par.sns.vc)){
  dat.name <- paste("SJI", par.sns[i, 3], "2023_02", par.sns[i, 1], sep = "_")
  dat.name <- paste(dat.name, "csv", sep = ".")
  dat <- sensors.raw[[dat.name]]
  dat.new <- data.frame(site = substr(dat.name, start = 1, stop = 5),
                       sensor_sn = substr(dat.name, start = 15, stop = 18),
                       time = paste(dat[7:nrow(dat), 2], dat[7:nrow(dat), 3], sep = " "),
                       par_uncal = dat[7:nrow(dat), 4])
  
  par.dat <- rbind(par.dat, dat.new)
}

par.dat <- par.dat %>% 
  mutate(time = as.POSIXct(time, format = "%d/%m/%Y %H:%M:%S")) %>% 
  mutate(par_uncal = as.numeric(par_uncal))

par.dat <- left_join(par.dat, site.names)
```

#### Depth
Note these data are not direct readings of depth, rather measures of absolute pressure in kPa. These data will need to be calibrated. Also, these sensor also report temperature.
```{r}
## Select depth metadata ##
depth.sns <- deploy.sji %>% 
  filter(type == "depth") %>% 
  select(sensor, site, subsite)

depth.sns.vc <- par.sns$sensor

## Loop over sensors to create master depth dataframe ##
depth.dat <- data.frame()

for(i in 1:length(depth.sns.vc)){
  dat.name <- paste("SJI", depth.sns[i, 3], "2023_02", depth.sns[i, 1], sep = "_")
  dat.name <- paste(dat.name, "csv", sep = ".")
  dat <- sensors.raw[[dat.name]]
  dat.new <- data.frame(site = substr(dat.name, start = 1, stop = 5),
                       sensor_sn = substr(dat.name, start = 15, stop = 22),
                       time = dat[3:nrow(dat), 2],
                       abs_pres_kPa = dat[3:nrow(dat), 3],
                       temp_c = dat[3:nrow(dat), 4])
  
  depth.dat <- rbind(depth.dat, dat.new)
}

depth.dat <- depth.dat %>% 
  mutate(time = strptime(time, format = "%m/%d/%y %I:%M:%S %p")) %>% 
  mutate(time = as.POSIXct(time, format = "%Y-%m-%d %H:%M:%S")) %>% 
  mutate(abs_pres_kPa = as.numeric(abs_pres_kPa)) %>% 
  mutate(temp_c = as.numeric(temp_c))
```

### Calibration and correction
Calibrate and correct data as needed. 

1. PAR correction?
2. Correct values for depth across sites. Esp. PAR. How does that work?
3. Fine tune date trimming
4. Derived values. Daily mean, median, hi/low, range, anomaly?
5. Correct for percentage drift on conductivity

#### PAR
Use calibration coefficients to correct raw data
```{r}
par.dat <- left_join(par.dat, par.coefs[, 3:4], by = "sensor_sn") %>%
  mutate(estimate = if_else(is.na(estimate), 1, estimate),
         calib = par_uncal * estimate,
         par_cal = par_uncal * estimate * 900 * 0.000001)
```

Correct for depth of the sensor. 
```{r}

```


### Clean up data

#### Trim to deployment date range
```{r}
## Deployment times ##
deploy.time <- deploy.sji %>% 
  select(c(type, site, subsite, sensor, launch.date, launch.time, deploy.date, deploy.time, retrieve.date, retrieve.time)) %>% 
  mutate(launch = paste(launch.date, launch.time, sep = " "),
         deploy = paste(deploy.date, deploy.time, sep = " "),
         retrieve = paste(retrieve.date, retrieve.time, sep = " ")) %>% 
  mutate(launch = as.POSIXct(launch, format = "%m/%d/%Y %H:%M:%S"),
         deploy = as.POSIXct(deploy, format = "%m/%d/%Y %H:%M:%S"),
         retrieve = as.POSIXct(retrieve, format = "%m/%d/%Y %H:%M:%S"))

## Trim ##
# Temperature #
temp.dat <- temp.dat %>% 
  inner_join(deploy.time[, c(4, 11:13)], by = c("sensor_sn" = "sensor")) %>% 
  filter(time >= deploy, time <= retrieve) %>% 
  select(c(site, sensor_sn, time, temp_c))

# Conductivity #
cond.dat <- cond.dat %>% 
  inner_join(deploy.time[, c(4, 11:13)], by = c("sensor_sn" = "sensor")) %>% 
  filter(time >= deploy, time <= retrieve) %>% 
  select(c(site, sensor_sn, time, spc_cond_uScm, sal_ppt, temp_c, lo_cond_uScm, hi_cond_uScm))

# PAR #
par.dat <- par.dat %>% 
  inner_join(deploy.time[, c(4, 11:13)], by = c("sensor_sn" = "sensor")) %>% 
  filter(time >= deploy, time <= retrieve) %>% 
  select(c(site, sensor_sn, time, par_uncal, estimate, calib, par_cal))

# Depth #
depth.dat <- depth.dat %>% 
  inner_join(deploy.time[, c(4, 11:13)], by = c("sensor_sn" = "sensor")) %>% 
  filter(time >= deploy, time <= retrieve) %>% 
  select(c(site, sensor_sn, time, abs_pres_kPa, temp_c))
```


#### Metadata
Site names
```{r}
## Temperature ##
temp.dat <- temp.dat %>%
  mutate(site.name = case_when(
         site == "SJI_1" ~ "Orcas E",
         site == "SJI_2" ~ "Orcas W",
         site == "SJI_3" ~ "Picnic Cove",
         site == "SJI_4" ~ "Reef Net Bay",
         site == "SJI_5" ~ "Jackson Beach",
         site == "SJI_6" ~ "4th of July",
         site == "SJI_7" ~ "Westcott Bay",
         site == "SJI_8" ~"Mitchell Bay")
         )

temp.dat$site.name <- factor(temp.dat$site.name, levels = c("Orcas E", "Orcas W",
                                                            "Mitchell Bay", "Westcott Bay",
                                                            "Reef Net Bay", "Picnic Cove",
                                                            "4th of July", "Jackson Beach"))

## Conductivity ##
cond.dat <- cond.dat %>%
  mutate(site.name = case_when(
         site == "SJI_1" ~ "Orcas E",
         site == "SJI_2" ~ "Orcas W",
         site == "SJI_3" ~ "Picnic Cove",
         site == "SJI_4" ~ "Reef Net Bay",
         site == "SJI_5" ~ "Jackson Beach",
         site == "SJI_6" ~ "4th of July",
         site == "SJI_7" ~ "Westcott Bay",
         site == "SJI_8" ~"Mitchell Bay")
         )

cond.dat$site.name <- factor(cond.dat$site.name, levels = c("Orcas E", "Orcas W",
                                                            "Mitchell Bay", "Westcott Bay",
                                                            "Reef Net Bay", "Picnic Cove",
                                                            "4th of July", "Jackson Beach"))

## PAR ##
par.dat <- par.dat %>% 
  mutate(site.name = case_when(
         site == "SJI_1" ~ "Orcas E",
         site == "SJI_2" ~ "Orcas W",
         site == "SJI_3" ~ "Picnic Cove",
         site == "SJI_4" ~ "Reef Net Bay",
         site == "SJI_5" ~ "Jackson Beach",
         site == "SJI_6" ~ "4th of July",
         site == "SJI_7" ~ "Westcott Bay",
         site == "SJI_8" ~"Mitchell Bay")
         )

par.dat$site.name <- factor(par.dat$site.name, levels = c("Orcas E", "Orcas W",
                                                            "Mitchell Bay", "Westcott Bay",
                                                            "Reef Net Bay", "Picnic Cove",
                                                            "4th of July", "Jackson Beach"))
```

### Exploratory plots
#### Temperature
```{r}
ggplot(temp.dat %>% 
         filter(time >= as.POSIXct("2023-05-25 00:00:00"), time <= as.POSIXct("2023-08-27 00:00:00"))) + 
  geom_point(aes(x = time, y = temp_c)) + 
  facet_wrap(~site, ncol = 2, nrow = 4)
```

#### Conductivity and Salinity

```{r}
## Salinity ##
ggplot(cond.dat) + 
  geom_point(aes(x = time, y = sal_ppt)) + 
  facet_wrap(~site.name, ncol = 2, nrow = 4)

## Spec Cond vs. temp ##
ggplot(cond.dat) +
  geom_point(aes(x = sal_ppt, y = temp_c)) +
  facet_wrap(~site.name, ncol = 2, nrow = 4)

ggplot(cond.dat %>% 
         filter(site.name == "Orcas E") %>% 
         filter(time >= as.POSIXct("2023-05-25 00:00:00"), time <= as.POSIXct("2023-08-30 00:00:00"))) +
  geom_point(aes(x = time, y = sal_ppt))

ggplot(cond.dat %>% 
         filter(site.name == "4th of July") %>% 
         filter(time >= as.POSIXct("2023-05-25 00:00:00"), time <= as.POSIXct("2023-08-30 00:00:00"))) +
  geom_point(aes(x = time, y = sal_ppt))
```

#### PAR
```{r}
ggplot(par.dat) + 
  geom_point(aes(x = time, y = par_cal)) + 
  facet_wrap(~site.name, ncol = 2, nrow = 4)
```

### Poster Summaries and plots

```{r}
pal <- c("#FFC20A", "#994F00", "#0C7BDC", "#40B0A6", "#E66100", "#5D3A9B", "#D35FB7", "#005AB5", "#D41159", "#454545")
```

#### Summarize

Compute daily mean and min max
```{r}
## Temperature ##
temp.daily <- temp.dat %>% 
  filter(time >= as.POSIXct("2023-05-25 00:00:00"), time <= as.POSIXct("2023-08-27 00:00:00")) %>% 
  mutate(date = as.Date(time)) %>% 
  group_by(site_name, date) %>% 
  summarize(avg_temp = mean(temp_c),
            sd_temp = sd(temp_c),
            min_temp = min(temp_c),
            max_temp = max(temp_c)) %>% 
  mutate(date = as.POSIXct(date, format = "%Y-%m-%d"))

temp.daily$site_name <- factor(temp.daily$site_name, levels = c("Ocras East", "Orcas West", "Picnic Cove", "Reef Net Bay", "Jackson Beach", "4th of July", "Westcott Bay", "Mitchell Bay"))

## Salinity ##
sal.daily <- cond.dat %>% 
  filter(time >= as.POSIXct("2023-05-25 00:00:00"), time <= as.POSIXct("2023-08-27 00:00:00")) %>% 
  mutate(date = as.Date(time)) %>% 
  group_by(site_name, date) %>% 
  summarize(avg_sal = mean(sal_ppt),
            sd_sal = sd(sal_ppt),
            min_sal = min(sal_ppt),
            max_sal = max(sal_ppt)) %>% 
  mutate(date = as.POSIXct(date, format = "%Y-%m-%d"))

sal.daily$site_name <- factor(sal.daily$site_name, levels = c("Ocras East", "Orcas West", "Picnic Cove", "Reef Net Bay", "Jackson Beach", "4th of July", "Westcott Bay", "Mitchell Bay"))

## PAR ##
par.daily <- par.dat %>% 
  filter(time >= as.POSIXct("2023-05-25 00:00:00"), time <= as.POSIXct("2023-08-27 00:00:00")) %>% 
  mutate(date = as.Date(time)) %>%
  mutate(tod = strftime(time, format="%H:%M:%S")) %>%
  mutate(tod = as.POSIXct(tod, format="%H:%M:%S")) %>% 
  filter(tod >= as.POSIXct("06:00:00", format="%H:%M:%S"), time <= as.POSIXct("21:00:00" , format="%H:%M:%S")) %>%
  group_by(site_name, date) %>% 
  summarize(avg_par = mean(par_cal),
            sd_par = sd(par_cal),
            min_par = min(par_cal),
            max_par = max(par_cal)) %>% 
  mutate(date = as.POSIXct(date, format = "%Y-%m-%d"))

par.daily$site_name <- factor(sal.daily$site_name, levels = c("Ocras East", "Orcas West", "Picnic Cove", "Reef Net Bay", "Jackson Beach", "4th of July", "Westcott Bay", "Mitchell Bay"))
```

#### Plot
```{r}
## Temperature ##
png(filename = "../CERF poster figs/Temp_Site.png", width = 10, height = 14, units = "in", res = 300)
ggplot(temp.daily) +
  geom_point(aes(x = date, y = avg_temp, color = site_name)) +
  scale_color_manual(values = pal) +
  scale_x_datetime(date_labels = "%b", limits = c(as.POSIXct("2023-05-25 00:00:00"), as.POSIXct("2023-09-01 00:00:000"))) +
  scale_y_continuous(limits = c(8, 18), breaks = seq(8, 18, by = 2)) +
  ylab("Daily mean temperature (°C)") +
  xlab("Date") +
  theme(legend.position = "none") +
  theme(text = element_text(size = 26)) +
  facet_wrap(~ site_name, ncol = 2, nrow = 4, scales = "free")
dev.off()

## Salinity ##
png(filename = "../CERF poster figs/Sal_Site.png", width = 10, height = 14, units = "in", res = 300)
ggplot(sal.daily) +
  geom_point(aes(x = date, y = avg_sal, color = site_name)) +
  scale_color_manual(values = pal) +
  scale_x_datetime(date_labels = "%b", limits = c(as.POSIXct("2023-05-25 00:00:00"), as.POSIXct("2023-09-01 00:00:000"))) +
  scale_y_continuous(limits = c(24, 36), breaks = seq(24, 36, by = 2)) +
  ylab("Daily mean salinity (ppt)") +
  xlab("Date") +
  theme(legend.position = "none") +
  theme(text = element_text(size = 26)) +
  facet_wrap(~ site_name, ncol = 2, nrow = 4, scales = "free")
dev.off()

## PAR ##
png(filename = "../CERF poster figs/PAR_Site.png", width = 10, height = 14, units = "in", res = 300)
ggplot(par.daily) +
  geom_point(aes(x = date, y = avg_par, color = site_name)) +
  scale_color_manual(values = pal) +
  scale_x_datetime(date_labels = "%b", limits = c(as.POSIXct("2023-05-25 00:00:00"), as.POSIXct("2023-09-01 00:00:000"))) +
  #scale_y_continuous(limits = c(24, 36), breaks = seq(24, 36, by = 2)) +
  ylab("Daily mean salinity (ppt)") +
  xlab("Date") +
  theme(legend.position = "none") +
  theme(text = element_text(size = 26)) +
  facet_wrap(~ site_name, ncol = 2, nrow = 4, scales = "free")
```

